Data Analysis Project 1: Peer to peer Loans
========================================================
  

## Processing
Dowloading the data
```{r loadData,cache=TRUE}
setwd("/Users/torniai2/Dropbox/coursera/data_analysis/data_Analysis_March11th")
#fileURL="https://spark-public.s3.amazonaws.com/dataanalysis/samsungData.rda"
#download.file(fileURL, destfile='./data/samsungData.rda', method="curl")
#dateDownloaded <- date()
#dateDownloaded
load("./data/samsungData.rda")
```

Let's have a look at the data. 
```{r}
summary(samsungData)
names(samsungData)
#head(samsungData)
sum(is.na(samsungData))
# Here I can say that the last two columns are describing the 

```
Too many variables, at least no NAs
Let's try to understnad the outcome:
```{r}
summary(as.factor(samsungData$activity))
```

I have 6 possible recognized actions.
Now let's build our datasets
Anyway how to identify the variables to use in my model?

```{r}
trainingset <- subset(samsungData, samsungData$subject==c(1,3,5,6))
testset <- subset(samsungData, samsungData$subject==c(27,28,29,30))
```

Need to get rid of the subject column and set as factors the outcomes.
```{r}
# drop the subject column in both train and test set
trainingset$subject <- NULL
testset$subject <- NULL
# HAve the activity as factor
testset$activity <- as.factor(testset$activity)
trainingset$activity <- as.factor(trainingset$activity)
summary(testset$activity)
# Here make sure that the names of the variables are checked (otheriwse I have an error)
testset <- data.frame(testset, check.names=TRUE)
trainingset <- data.frame(trainingset, check.names=TRUE)
```
We see that the test set has more or less a uniform distribution across user's activities
It would be interesting to see across the 4 users how is this distribution 

## Prediction 

Now Let's set the seed and build a random forest
```{r}
library(randomForest)
library(ElemStatLearn)
library(e1071)
#Set the seed
set.seed(271188)
# Now random Forest
forestmodel <- randomForest(activity ~., data=testset, prox=TRUE)
forestmodel
```

Now let's see how this simple model works on predicition
```{r}
pred <- predict(forestmodel, testset)
# Let's evaluate the error
error = sum(pred!=testset$activity)/length(testset$activity)
# Impossibile mi viene errore zero... ma qualcosa non va oviamente.
# Per adesso te lo lascio qua.

# Andra' fatto un set do forest tree
```

summary(testes$activity)