Lifesaver Accelerometer Test
========================================================

A few tests done on accelerometers for the [Lifesaver Project](https://life-saver.org.ok).

```{r}
# 0001.csv contains all PUSH detections
# con <- file('./data/0001.csv', 'r')
# accData0001 <- read.csv(con)

# 0002.csv is same as 0001 but PUSH has been removed
# con <- file('./data/0002.csv', 'r')
# accData0002 <- read.csv(con)

# 0003.csv has a casual cpr at the beginning, then goes crazy
# sample rate is still every 10ms and norms > 9 are boosted with a x3 factor
# con <- file('./data/0003.csv', 'r')
# accData0003 <- read.csv(con)

con <- file('./data/0004.csv', 'r')
accData0004 <- read.csv(con)

close(con)
```

* accData0002 is same accelerometer data as accData0001 where PUSH actions have been removed.
* accData0003 has a casual cpr at the beginning, then goes crazy. sample rate is still every 10ms and norms > 9 are boosted with a x3 factor
* accData0004 traces y with history, data is pre-processed so as to remove gravity

Now we have the data, start with a bit of exploration and create local variables

```{r}
sapply(accData0004[1,], class)
head(accData0004)
summary(accData0004)

procAccData <- accData0004
```

procAccData is a subset of accData002 from frame 300 since it's the moment when the relevant CPR action starts. OK, now plot frequency and and scatter.

```{r fig.width=7, fig.height=6}
hist(sample(procAccData[,1],300))
plot(data.frame(cbind(1:length(procAccData[,1]), procAccData[,1])), pch=20, cex=0.4, col="#2222AA", type="o", xlab="frames", ylab="y-acc")
plot(data.frame(cbind(1:length(procAccData[,2]), procAccData[,2])), pch=20, cex=0.4, col="#AA2222", type="o", xlab="frames", ylab="y-acc history")
```

Main and history plots for Y-acceleration look the same, although peak values are lower

```{r}
matplot(procAccData[800:1100,], type="l", col=c("#2222AA", "#AA2222"))
```

```{r}
# fuck around with clusters
distMatrix <- dist(sample(procAccData[200:800,1], 200), sample(procAccData[800:1400,2], 200), method="euclidean")
hClust <- hclust(distMatrix)
plot(hClust)
```
